---
title: "SGControl"
output: html_document
date: "2023-07-18"
---
SG Personal Pipeline for Dinimite Cruise June 2022 Metabolite Data.

Goals:

Section I: Import and cleaning/rearranging of data.
Section II: Quality control using user-defined parameters. 
Section III: Applying Best-Matched Internal Standard (B-MIS).
Section IV: Quantifying peak area to umol/vial when possible.


Load your libraries
```{r setup, include=TRUE}
library(tidyverse)
library(janitor)
```

# Section I: Import and cleaning/rearranging of data.


Import the file that came directly from Skyline. Assign it a variable name and save it to the directory.Rearrange the dataset into a file you like: Compound name, Filename, Area.
```{r setup, include=TRUE}
#Creating the filepath for file
filepath <- "data_raw/Dino22HILIPos.csv"

#Having an original file of the file received directly from Skyline
ogHILICpos <- read.csv(filepath, check.names = TRUE) %>%
  clean_names()

#Reading the file into the directory and clearing the columns you don't want
HILICpos <- read.csv(filepath, check.names = TRUE) %>%
  clean_names() %>%
  mutate(filename = replicate_name,
         compound_name = precursor_ion_name,
         area = as.numeric(area),
         retention_time = as.numeric(retention_time),
         height = as.numeric(height),
         background = as.numeric(background)) %>%
  select("compound_name", "filename", "area", "retention_time", "height", "background") 

#Below this can check for multiple entries per compoundname and filename
#  count(compound_name, filename)
```
Import Ingalls Standards List, clean and arrange with columns you desire. Here the compound name and the retention times are important for me.

```{r setup, include=TRUE}
stds_list <- read.csv("data_raw/Ingalls_Lab_Standards.csv", check.names = TRUE) %>%
  clean_names() %>%
  select("compound_name", "rt_minute")
```


# Section II: Quality control using user-defined parameters.

Setting the parameters desired. 
```{r setup, include=TRUE}
# QE QC parameters
area.min   <- 1000 # HILIC - 1000
rt.flex    <- 3 # HILIC +/- 0.4 min
blk.thresh <- 0.3 # HILIC +/- 0.3
sn.min     <- 4 # HILIC - 4
height.min <- 1000
```

Joining dataframes in order to create dataframe that can change set up the parameters. 
```{r}
wip_QC_HILICpos <- inner_join(HILICpos, stds_list, by = "compound_name") %>%
  mutate(retention_time = as.numeric(retention_time),
         std_retention_time = as.numeric(rt_minute),
         qc_area = area> area.min,
         qc_abs_rt = abs(rt_minute - retention_time),
         qc_rt = qc_abs_rt > rt.flex,
         sn = (area / background),
         qc_sn = sn > sn.min,
         qc_height = height > height.min)
```

Creating Blank Dataframe
```{r}
 blank_table <- HILICpos %>%
  filter(str_detect(filename, "Blk")) %>%
  group_by(compound_name) %>%
  mutate(blank_max = max(area)) %>%
  select(compound_name, blank_max) %>%
  unique()
```

Joining Blank Dataframe and wip_QC_HILICpos.
```{r}
QC_HILICpos <- inner_join(wip_QC_HILICpos, blank_table, by = "compound_name") %>%
  mutate(qc_blk_thresh = area > blank_max) %>%
  select("compound_name", "filename", "area", "qc_area", "retention_time","qc_rt", "height","qc_height", "sn", "qc_sn", "qc_blk_thresh", "background")
  
```

# Section III: Applying Best-Matched Internal Standard (B-MIS)

1. Make dataframe into wide format
```{r}
BMIS_df <- ogHILICpos %>%
  clean_names() %>%
  mutate(filename = replicate_name,
         compound_name = precursor_ion_name) %>%
  select( "filename", "compound_name", "area")

# making it into a wide format
BMIS_df_wide <- BMIS_df %>%
  pivot_wider(
    id_cols = c("filename"),
    names_from = "compound_name",
    values_from = "area")
```

2.Create a new dataframe that only contains the Internal Standards
```{r}
# Here I am listing all the internal standards I have in this sheet
IS <- c("Adenine, 15N2","Arsenobetaine, 13C2", "Cytosine, 13C2, 15N3", "DL-Alanine, 2H3", "DL-Histidine, 15N", "DL-Proline, 2H7", "DL-Valine, 2H8","Glycine betaine, 13C5, 15N", "Guanine, 13C, 15N2", "Homarine, 2H3", "L-Isoleucine, 15N", "L-Methionine, 2H3", "Trimethylamine N-oxide, 2H9","L-Arginine, 13C6, 15N4")

IS_BMIS_df <- BMIS_df_wide %>%
  select(filename, all_of(IS))

# Now you will filter for only the "Poo"

# setting the name of the string I am interested
poostring <- "221006_Poo_TruePooRC_"

# Filter rows for filenames containing the specified string
IS_poo_BMIS_df <- IS_BMIS_df %>%
  filter(str_detect(filename, poostring)) %>%
  slice(-(1:3)) %>%

```

3. Normalize BMIS areas and gather CV for compound. 
```{r}
# Creating a datafram that only contains the poo with all the compounds and no standards
poo_cmpds_BMIS_df <- BMIS_df_wide %>%
  filter(str_detect(filename, poostring)) %>%
  slice(-(1:3)) %>%
  select(-all_of(IS))

# Finding Normalizes Peak Areas

BMIS_df_wide$Adenine <- as.numeric(BMIS_df_wide$Adenine)
BMIS_df_wide$`Adenine, 15N2` <- as.numeric(BMIS_df_wide$`Adenine, 15N2`)

poo_BMIS_df <- BMIS_df_wide %>%
  filter(str_detect(filename, poostring)) %>%
  slice(-(1:3)) %>%
  mutate(BMIS_adenine = (Adenine / `Adenine, 15N2`))




```

